---
title: "Run to equilibrium and investigate a school age children mda program"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run to equilibrium and investigate a school age children mda program}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
First we have to load the needed library:
```{r setup}
library(SchistoxpkgR)
```

After this, we will need to setup the environment to use Julia and load the required Julia libraries by running
```{r setup2}
schistox_setup()
```

# Define and initialize parameters
As we have many parameters, we will then need to define these so that we can run simulations. As these are passed to Julia to run, some parameters require special formatting, such as population size must be passed ```as.integer(N)```. Below we define all parameters required and initialize the ```Parameters``` object needed to run the code. Definitions of the parameters can be found by running ```?set_pars```.
```{r define_params}
N = as.integer(1000)    
time_step = 10
N_communities = 1
community_probs = JuliaCall::julia_eval("[1.0]", need_return =  "Julia")
community_contact_rate = JuliaCall::julia_eval("[1.0]", need_return =  "Julia")
density_dependent_fecundity = 0.0007 # for S. mansoni [Toor et al JID paper SI]
#density_dependent_fecundity = 0.0006 # for S. haematobium [Toor et al JID paper SI]
average_worm_lifespan = 5.7 # years for S. mansoni [Toor et al JID paper SI]
#average_worm_lifespan = 4 # years for S. haematobium [Toor et al JID paper SI]
max_age = 100
initial_worms = 0

# miracidia_maturity = 21 # for S. haemotobium 
miracidia_maturity = 24 # for S. mansoni 
initial_miracidia = 100000*N/1000
initial_miracidia_days = round(miracidia_maturity/time_step)
init_env_cercariae = 100000*N/1000
worm_stages = as.integer(1)
contact_rate = 0.09
max_fecundity = 0.34  # for S. mansoni [Toor et al JID paper SI]
#max_fecundity = 0.3  # for S. haematobium [Toor et al JID paper SI]
age_contact_rates = c(0.032, 0.610, 1, 0.06)
ages_for_contacts = c(4, 9, 15, 100)
contact_rate_by_age_array = rep(0,101)

mda_adherence = .9
mda_access = .9
female_factor = 1
male_factor = 1
birth_rate = 28*time_step/(1000*365)
human_cercariae_prop = 1

predis_aggregation = 0.24 # 0.24 for high prev settings; 0.04 for low prev settings # From "The design of schistosomiasis monitoring and evaluation programmes:
#The importance of collecting adult data to inform treatment strategies for Schistosoma mansoni"
cercariae_survival = 1/2
miracidia_survival = 1/2
death_prob_by_age = c(0.0656, 0.0093, 0.003, 0.0023, 0.0027, 0.0038, 0.0044, 0.0048, 
                      0.0053, 0.0065, 0.0088, 0.0106, 0.0144, 0.021, 0.0333, 0.0529, 
                      0.0851, 0.1366, 0.2183, 0.2998 , 0.3698, 1)
ages_for_death = c(1, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60,
                   65, 70, 75, 80, 85, 90, 95, 100, 110)
r = 0.03
vaccine_effectiveness = 0.95
drug_effectiveness = 0.863
spec_ages = c(8639, 9082, 6424, 5074, 4425, 3847, 3628, 3062,
              2436, 1770, 1868, 1066, 743, 518, 355, 144)
ages_per_index = 5
record_frequency = 1/24
use_kato_katz = 0
kato_katz_par = 0.87
heavy_burden_threshold = 16
rate_acquired_immunity = 0
M0 = 20
human_larvae_maturity_time = 30
input_ages = array(0,dim=c(0))
input_contact_rates = array(0,dim=c(0))
scenario = "moderate adult"

pars = set_pars(N,
                time_step,
                N_communities,
                community_probs,
                community_contact_rate,
                density_dependent_fecundity,
                average_worm_lifespan,
                max_age,
                initial_worms,
                initial_miracidia,
                initial_miracidia_days,
                init_env_cercariae,
                worm_stages,
                contact_rate,
                max_fecundity,
                age_contact_rates,
                ages_for_contacts,
                contact_rate_by_age_array,
                mda_adherence,
                mda_access,
                female_factor,
                male_factor,
                miracidia_maturity,
                birth_rate,
                human_cercariae_prop,
                predis_aggregation,
                cercariae_survival,
                miracidia_survival,
                death_prob_by_age,
                ages_for_death,
                r,
                vaccine_effectiveness,
                drug_effectiveness,
                spec_ages,
                ages_per_index,
                record_frequency,
                use_kato_katz,
                kato_katz_par,
                heavy_burden_threshold,
                rate_acquired_immunity,
                M0,
                human_larvae_maturity_time,
                input_ages, 
                input_contact_rates, 
                scenario)

```

# Generate population and simulate to equilibrium
Once we have defined the parameters, we can create the population and begin running simulations. We must define mda and vaccine information here. If we wish to perform no MDA or vaccinations then we add these as ```mda_info = array(0,dim=c(0,0)); vaccine_info = array(0,dim=c(0,0))``` as seen below.:
```{r create_population}
e = create_population_specified_ages(pars)
humans = generate_ages_and_deaths(20000, e$humans, pars)
humans = update_contact_rate(humans,  pars)


mda_info = array(0,dim=c(0,0))
vaccine_info = array(0,dim=c(0,0))
number_years_equ = 100
num_time_steps = as.integer(365*number_years_equ / time_step)
num_repeats = 1



e = update_env_constant_population_human_larvae(num_time_steps, humans,  e$miracidia, e$cercariae, pars, mda_info, vaccine_info)
d = get_data_from_record(e$record)

sac_burden = d[[1]]
sac_heavy_burden = d[[2]]
times = d[[3]]
```




``` {r plot, echo = FALSE , fig.dim = c(8, 6)}
plot(times, sac_burden, type = 'l', col = rgb(110/255, 99/255, 252/255), bty = 'n',
     ylim = c(0, max(sac_burden)), lwd = 2, xlab = "year", ylab = "SAC prevalence")
lines(times, sac_heavy_burden, col = rgb(30/255, 190/255, 160/255), lwd =2 )
abline(h = 1, lwd = 2, lty = 2, col = rgb(2/255, 163/255, 217/255))
abline(h=c(0,10,20, 30,40,50,60,70,80,90), col = 'lightgrey')
legend('topright',legend=c("SAC prev", "SAC heavy burden"),
       col=c(rgb(110/255, 99/255, 252/255), rgb(30/255, 190/255, 160/255)), lwd = c(2,2), lty = c(1,1), cex=1.2,
       title="", text.font=18, bg='lightblue', bty = 'n')
```

We can save the resulting population as follows:
```{r save_pop}
filename = "population.jld"
save_population_to_file(filename, e$humans,  e$miracidia, e$cercariae, pars)
```

# Consider effect of MDA
We can then investigate the use of MDA's in reducing the prevalence of schistosomiasis in the population. The MDA performed here is 75% coverage in SAC once a year and is defined in the ```create_mda``` function (run ```?create_mda``` for docs):
```{r mda}
num_repeats = 10 #number of simulations to run
number_years = 20
drug_efficacy = 0.863 #Toor et al. JID paper in SI: drug efficacy 86.3% for S. mansoni and 94% for S. haematobium
num_time_steps = as.integer(365*number_years / time_step)

mda_info = create_mda(0, .75, 0, 1,
                      number_years, 1, c(0,1), c(0,1), c(0,1), drug_efficacy)



e1 =  run_repeated_sims_no_population_change_human_larvae(filename, num_time_steps, mda_info, vaccine_info, num_repeats)




```

```{r, echo = FALSE, fig.dim = c(8, 6)}
plot(e1$times, get_dot_mean(e1$sac_prev), type = 'l', col = rgb(110/255, 99/255, 252/255), bty = 'n',
     ylim = c(0,  max(get_dot_mean(e1$sac_prev))), lwd = 2, xlab = "year", ylab = "SAC prevalence")
lines(e1$times, get_dot_mean(e1$high_burden_sac), col = rgb(30/255, 190/255, 160/255), lwd =2 )
abline(h = 1, lwd = 2, lty = 2, col = rgb(2/255, 163/255, 217/255))
abline(v=c(0,5,10,15,20), col = 'lightgrey')
abline(h=c(0,10,20, 30,40,50,60,70,80,90), col = 'lightgrey')
legend('topright',legend=c("SAC prev", "SAC heavy burden",
                           'missed MDA window', 'heavy burden goal'),
       col=c(rgb(110/255, 99/255, 252/255), rgb(30/255, 190/255, 160/255)
             ,'black', rgb(2/255, 163/255, 217/255)), lwd = c(2,2,2, 2), 
       lty = c(1,1,2,2), cex=1.2,
       title="", text.font=18, bg='lightblue', bty = 'n')

```

# Run individual functions

We may wish to run some of these functions in isolation, for example to perform an MDA on the population once time and then run the simulation forward again, or to change the parameters of the simulation and then continue a simulation.

To perform an MDA of 80% of the population we can run:
```{r individual_mda}
humans = administer_drug(e$humans, indices = sample(1:N, N * .8), drug_effectiveness = 0.863)
```
Note that to perform a more sophisticated MDA we must input the indices of the MDA more carefully, for example finding all individuals whose age is in a certain range first, and then specifying the proportion who will receive MDA.

We can also update the contact rate by age in the following way:
```{r update_contact_rate_by_age}
input_ages = array(data = c(as.integer(4),as.integer(9),as.integer(15), as.integer(max_age)))
input_contact_rates = array(data = c(0.05,	0.25,	0.62,	0.1))
input_contact_rates = input_contact_rates/sum(input_contact_rates)
pars = make_age_contact_rate_array(pars, scenario, input_ages, input_contact_rates)
```
```input_ages``` specifies the age groups which we are specifying the contact rates for. ```input_contact_rate``` are the chosen contact rates for the age groups specified. Having a 4 in the in the first position means that individuals from the ages of 0 to 4 will have the contact rate given by the first entry in ```input_contact_rate```. From 5 to the next specified age, individuals in this age range will have the 2nd specified contact rate and so on. Here the contact rates are normalized so that the sum of the contact rate array is 1, though this isn't necessary. The 4th line here is where the contact rates are actually updated in the parameters object.

We can also update individual parameters of the simulation using the ```update_parameters_individually``` function. This is used as follows:
```{r update_pars}
update_parameters_individually(pars, name = "max_fecundity", value = 0.32)
```
where the ```name``` input must exactly match a parameter name in the ```Parameter``` variable and ```value``` is the updated parameter value. This can be used to update the global contact rate, which could be used if we wanted to simulate a WASH intervention or the building or upgrading of latrines.


To return every entry of a chosen variable from the array containing the human population data, which is a Julia array, we run the following:
```{r return_variable}
eggs = get_selected_data(humans, "eggs")
ages = get_selected_data(humans, "age")
female_worms = get_selected_data(humans, "female_worms")
male_worms = get_selected_data(humans, "male_worms")
```

The 2nd input in this function must be exactly the name of a variable in the humans object. Options for this are:
```"age", "death_age", "gender", "predisposition", "female_worms", "male_worms", "eggs", "vac_status", "age_contact_rate", "adherence", "access", "community", "relative_contact_rate", "uptake_rate", "acquired_immunity", "total_worms", "larvae"```.

